---
title: "beeSpec"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#knitr::opts_chunk$set(fig.width=3, fig.height=5) # use only to create multipanel plot
```

# load packages
```{r}
library(tidyverse)
library(dplyr)
library(plyr)
library(ggplot2)
library(factoextra)
library(vegan) 
library(lme4)
library(plotly)
library(multcomp) # glht tukey
library(sjPlot)   # tab_model
library(ggpubr)   # ggdensity
library(effects)  # plot output of glmm
library(MuMIn)    # model.sel
```

# load data
```{r}
# pooled data across timepoints, values are counts
# includes all bees, but only known bees with >=3 observations have shannon values
count_data <- read.csv("../Data/filtData/count_data.csv") 


count_replac <- count_data[count_data$beeType == "replac",]
count_replac <- count_replac[!is.na(count_replac$beeID),]
count_replac

count_orig <- count_data[count_data$beeType == "orig",]
count_orig <- count_orig[!is.na(count_orig$beeID),]
count_orig

ggplot(count_data, aes(x = nBehavs, y = beeType)) +
  geom_point()

mean(count_replac$nBehavs)
sd(count_replac$nBehavs)/sqrt(nrow(count_replac))

mean(count_orig$nBehavs)
sd(count_orig$nBehavs)/sqrt(nrow(count_orig))

#How many included?
count_data %>% filter(lazy=="notLazy") %>% nrow()

#How many not included, lazy? 
count_data %>% filter(lazy=="lazy") %>% nrow()


```

# count number of specialists
```{r}

nrow(count_data[count_data$specBehav == "nectarforaging",])
nrow(count_data[count_data$specBehav == "pollenforaging",])
nrow(count_data[count_data$specBehav == "feedsbrood",])
nrow(count_data[count_data$specBehav == "egglaying",])
nrow(count_data[count_data$specBehav == "generalist",])
nrow(count_data[count_data$specBehav == "lazy",])

```

# visualize body size in relation to behavioral specialization
```{r}
count_data$specBehav <- ordered(count_data$specBehav, levels = c("lazy", "generalist", "nectarforaging", "feedsbrood", "egglaying"))
p1 <- ggplot(data = count_data, aes(size, specBehav, color = shannon, shape = caste)) +
  geom_point(position = position_jitter(width=0.1, height=0.2)) +
  stat_summary(fun = mean, geom = "point", size = 5) +
  scale_shape_manual(values = c(1, 16)) +
  scale_color_gradient(low="dark blue", high="yellow") +
  # scale_color_manual(values = c("#F8766D", "#7CAE00", "#00BFC4", "#8c8c8c", "#ffc425")) +
  labs(x = "body size", y = "behavioral specialization", color = "shannon")
p1
ggsave("../figures/bodySize.jpg")
```

# visualize ovary development in relation to behavioral specialization
```{r}
# OVARY LENGTH
ggplot(data = count_data, aes(ovLavg, specBehav, color = shannon)) +
  geom_point(position = position_jitter(width=0.1, height=0.2), alpha = 0.8) +
  stat_summary(fun = mean, geom = "point", size = 5) +
  scale_color_gradient(low="dark blue", high="yellow") +
  # scale_color_manual(values = c("#F8766D", "#7CAE00", "#00BFC4", "grey", "black")) +
  labs(x = "ovary length", y = "behavioral specialization", color = "behavioral specialization")
ggsave("../figures/ovaryLength.jpg")


# OVARY STAGE (resorbed oocytes = stage 4)
p2 <- ggplot(data = count_data, aes(ovSavg, specBehav, color = shannon)) +
  geom_point(position = position_jitter(width=0.1, height=0.2), alpha = 0.8) +
  stat_summary(fun = mean, geom = "point", size = 5) +
  scale_color_gradient(low="dark blue", high="yellow") +
  # scale_color_manual(values = c("#F8766D", "#7CAE00", "#00BFC4", "#8c8c8c", "#ffc425")) +
  labs(x = "ovary stage", y = "behavioral specialization", color = "shannon")
p2
ggsave("../figures/ovaryStage.jpg")


# OOCYTE RESORPTION - similar trends
p3 <- ggplot(data = count_data, aes(ovRavg, specBehav, color = shannon)) +
  geom_point(position = position_jitter(width=0.1, height=0.2), alpha = 0.8) +
  stat_summary(fun = mean, geom = "point", size = 5) +
  scale_color_gradient(low="dark blue", high="yellow") +
  # scale_color_manual(values = c("#F8766D", "#7CAE00", "#00BFC4", "#8c8c8c", "#ffc425")) +
  labs(x = "resorption status", y = "behavioral specialization", color = "shannon")
p3
ggsave("../figures/ovarResorb.jpg")
```

```{r}
ggarrange(p1, p2, p3, nrow = 3, labels = c("A", "B", "C"))
ggsave("../figures/behavSpecBSov.jpg")
```

# glmms on behavioral specialization vs. oocyte length
```{r}
# it's not really normal... not sure what distribution to use
ggdensity(count_data$size)
count_data

# transform the data to make each specialization a different column, with yes/no (0/1) data in each row
count_wide <- pivot_wider(count_data, id_cols = c(caste, nestTYPE, beeID, nestID, size, ovSavg, ovRavg, natal, queen, nBees), names_from = specBehav, values_from = specBehav)

# make NA values "other" in behavioral specialization
count_wide$feedsbrood <- as.character(count_wide$feedsbrood)
count_wide[is.na(count_wide$feedsbrood), "feedsbrood"] <- "other"
count_wide$feedsbrood <- factor(count_wide$feedsbrood, levels = c("feedsbrood", "other"))

count_wide$nectarforaging <- as.character(count_wide$nectarforaging)
count_wide[is.na(count_wide$nectarforaging), "nectarforaging"] <- "other"
count_wide$nectarforaging <- factor(count_wide$nectarforaging, levels = c("nectarforaging", "other"))

count_wide$egglaying <- as.character(count_wide$egglaying)
count_wide[is.na(count_wide$egglaying), "egglaying"] <- "other"
count_wide$egglaying <- factor(count_wide$egglaying, levels = c("egglaying", "other"))

count_wide$lazy <- as.character(count_wide$lazy)
count_wide[is.na(count_wide$lazy), "lazy"] <- "other"
count_wide$lazy <- factor(count_wide$lazy, levels = c("lazy", "other"))

count_wide$generalist <- as.character(count_wide$generalist)
count_wide[is.na(count_wide$generalist), "generalist"] <- "other"
count_wide$generalist <- factor(count_wide$generalist, levels = c("generalist", "other"))

count_wide$caste <- factor(count_wide$caste, levels = c("Q", "W"))
count_wide
# need to add natal colony as random effect for all these analyses
```

## egg laying
```{r}
# MODEL EGG LAYING 
el0 <- glmer(egglaying ~ 1 + (1|natal) + (1|nestID),
              data = count_wide, family = binomial)
el1 <- glmer(egglaying ~ size + ovSavg + ovRavg + queen + nBees + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
el2 <- glmer(egglaying ~ size + ovSavg + ovRavg + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
el3 <- glmer(egglaying ~ size + ovSavg + queen + nBees + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
el4 <- glmer(egglaying ~ size + ovSavg + queen  + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
el5 <- glmer(egglaying ~ size + ovSavg  + nBees + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
el6 <- glmer(egglaying ~ size + ovRavg + queen + nBees + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
el7 <- glmer(egglaying ~ size + ovRavg + queen  + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
el8 <- glmer(egglaying ~ size + ovRavg  + nBees + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
el9 <- glmer(egglaying ~ ovSavg + ovRavg + queen + nBees + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
el10 <- glmer(egglaying ~ ovSavg + ovRavg + queen  + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
el11 <- glmer(egglaying ~ ovSavg + ovRavg  + nBees + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
el12 <- glmer(egglaying ~ size + ovSavg + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
el13 <- glmer(egglaying ~ size + ovRavg + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
el14 <- glmer(egglaying ~ ovSavg + ovRavg + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
el15 <- glmer(egglaying ~ size + queen + nBees + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
el16 <- glmer(egglaying ~ size + queen  + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
el17 <- glmer(egglaying ~ size  + nBees + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
el18 <- glmer(egglaying ~ ovSavg + queen + nBees + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
el19 <- glmer(egglaying ~ ovSavg + queen  + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
el20 <- glmer(egglaying ~ ovSavg  + nBees + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
el21 <- glmer(egglaying ~ ovRavg + queen + nBees + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
el22 <- glmer(egglaying ~ ovRavg + queen  + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
el23 <- glmer(egglaying ~ ovRavg  + nBees + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
el24 <- glmer(egglaying ~ size + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
el25 <- glmer(egglaying ~ ovSavg + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
el26 <- glmer(egglaying ~ ovRavg + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
el27 <- glmer(egglaying ~ queen + nBees + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
el28 <- glmer(egglaying ~ queen  + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
el29 <- glmer(egglaying ~ nBees + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
model.sel(el0, el2, el5, el6, el7, el8, el10, el11, el13, el14, el15, el16, el17, el18, el19, el20, el21, el22, el23, el24, el25, el26, el27, el28, el29)
mymodel <- el24
summary(mymodel)
tab_model(mymodel)
plot(allEffects(mymodel))
```

## brood feeding
```{r}
# MODEL BROOD FEEDING
fb0 <- glmer(feedsbrood ~ 1 + (1|natal) + (1|nestID) + (1|caste),
              data = count_wide, family = binomial)
fb1 <- glmer(feedsbrood ~ size + ovSavg + ovRavg + queen + nBees + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
fb2 <- glmer(feedsbrood ~ size + ovSavg + ovRavg + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
fb3 <- glmer(feedsbrood ~ size + ovSavg + queen + nBees + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
fb4 <- glmer(feedsbrood ~ size + ovSavg + queen  + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
fb5 <- glmer(feedsbrood ~ size + ovSavg  + nBees + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
fb6 <- glmer(feedsbrood ~ size + ovRavg + queen + nBees + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
fb7 <- glmer(feedsbrood ~ size + ovRavg + queen  + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
fb8 <- glmer(feedsbrood ~ size + ovRavg  + nBees + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
fb9 <- glmer(feedsbrood ~ ovSavg + ovRavg + queen + nBees + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
fb10 <- glmer(feedsbrood ~ ovSavg + ovRavg + queen  + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
fb11 <- glmer(feedsbrood ~ ovSavg + ovRavg  + nBees + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
fb12 <- glmer(feedsbrood ~ size + ovSavg + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
fb13 <- glmer(feedsbrood ~ size + ovRavg + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
fb14 <- glmer(feedsbrood ~ ovSavg + ovRavg + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
fb15 <- glmer(feedsbrood ~ size + queen + nBees + (1|nestID) + (1|natal),
              data = count_wide, family = binomial)
fb16 <- glmer(feedsbrood ~ size + queen  + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
fb17 <- glmer(feedsbrood ~ size  + nBees + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
fb18 <- glmer(feedsbrood ~ ovSavg + queen + nBees + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
fb19 <- glmer(feedsbrood ~ ovSavg + queen  + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
fb20 <- glmer(feedsbrood ~ ovSavg  + nBees + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
fb21 <- glmer(feedsbrood ~ ovRavg + queen + nBees + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
fb22 <- glmer(feedsbrood ~ ovRavg + queen  + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
fb23 <- glmer(feedsbrood ~ ovRavg  + nBees + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
fb24 <- glmer(feedsbrood ~ size + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
fb25 <- glmer(feedsbrood ~ ovSavg + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
fb26 <- glmer(feedsbrood ~ ovRavg + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
fb27 <- glmer(feedsbrood ~ queen + nBees + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
fb28 <- glmer(feedsbrood ~ queen  + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
fb29 <- glmer(feedsbrood ~ nBees + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
model.sel(fb0, fb2, fb5, fb6, fb7, fb8, fb10, fb11, fb13, fb14, fb15, fb16, fb17, fb18, fb19, fb20, fb21, fb22, fb23, fb24, fb25, fb26, fb27, fb28, fb29)

mymodel <- fb24
summary(mymodel)
tab_model(mymodel)
plot(allEffects(mymodel))
```
## nectar foraging
```{r}
# MODEL NECTAR FORAGING
nf0 <- glmer(nectarforaging ~ 1 + (1|natal) + (1|nestID),
              data = count_wide, family = binomial)
nf1 <- glmer(nectarforaging ~ size + ovSavg + ovRavg + queen + nBees + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
nf2 <- glmer(nectarforaging ~ size + ovSavg + ovRavg + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
nf3 <- glmer(nectarforaging ~ size + ovSavg + queen + nBees + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
nf4 <- glmer(nectarforaging ~ size + ovSavg + queen  + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
nf5 <- glmer(nectarforaging ~ size + ovSavg  + nBees + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
nf6 <- glmer(nectarforaging ~ size + ovRavg + queen + nBees + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
nf7 <- glmer(nectarforaging ~ size + ovRavg + queen  + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
nf8 <- glmer(nectarforaging ~ size + ovRavg  + nBees + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
nf9 <- glmer(nectarforaging ~ ovSavg + ovRavg + queen + nBees + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
nf10 <- glmer(nectarforaging ~ ovSavg + ovRavg + queen  + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
nf11 <- glmer(nectarforaging ~ ovSavg + ovRavg  + nBees + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
nf12 <- glmer(nectarforaging ~ size + ovSavg + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
nf13 <- glmer(nectarforaging ~ size + ovRavg + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
nf14 <- glmer(nectarforaging ~ ovSavg + ovRavg + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
nf15 <- glmer(nectarforaging ~ size + queen + nBees + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
nf16 <- glmer(nectarforaging ~ size + queen  + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
nf17 <- glmer(nectarforaging ~ size  + nBees + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
nf18 <- glmer(nectarforaging ~ ovSavg + queen + nBees + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
nf19 <- glmer(nectarforaging ~ ovSavg + queen  + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
nf20 <- glmer(nectarforaging ~ ovSavg  + nBees + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
nf21 <- glmer(nectarforaging ~ ovRavg + queen + nBees + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
nf22 <- glmer(nectarforaging ~ ovRavg + queen  + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
nf23 <- glmer(nectarforaging ~ ovRavg  + nBees + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
nf24 <- glmer(nectarforaging ~ size + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
nf25 <- glmer(nectarforaging ~ ovSavg + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
nf26 <- glmer(nectarforaging ~ ovRavg + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
nf27 <- glmer(nectarforaging ~ queen + nBees + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
nf28 <- glmer(nectarforaging ~ queen  + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
nf29 <- glmer(nectarforaging ~ nBees + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
model.sel(nf0, nf2, nf5, nf6, nf7, nf8, nf10, nf11, nf13, nf14, nf15, nf16, nf17, nf18, nf19, nf20, nf21, nf22, nf23, nf24, nf25, nf26, nf27, nf28, nf29)

mymodel <- nf24
summary(mymodel)
tab_model(mymodel)
plot(allEffects(mymodel))
```

## generalists
```{r}
# MODEL GENERALISTS
gen0 <- glmer(generalist ~ 1 + (1|natal) + (1|nestID),
              data = count_wide, family = binomial)
gen1 <- glmer(generalist ~ size + ovSavg + ovRavg + queen + nBees + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
gen2 <- glmer(generalist ~ size + ovSavg + ovRavg + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
gen3 <- glmer(generalist ~ size + ovSavg + queen + nBees + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
gen4 <- glmer(generalist ~ size + ovSavg + queen  + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
gen5 <- glmer(generalist ~ size + ovSavg  + nBees + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
gen6 <- glmer(generalist ~ size + ovRavg + queen + nBees + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
gen7 <- glmer(generalist ~ size + ovRavg + queen  + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
gen8 <- glmer(generalist ~ size + ovRavg  + nBees + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
gen9 <- glmer(generalist ~ ovSavg + ovRavg + queen + nBees + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
gen10 <- glmer(generalist ~ ovSavg + ovRavg + queen  + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
gen11 <- glmer(generalist ~ ovSavg + ovRavg  + nBees + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
gen12 <- glmer(generalist ~ size + ovSavg + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
gen13 <- glmer(generalist ~ size + ovRavg + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
gen14 <- glmer(generalist ~ ovSavg + ovRavg + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
gen15 <- glmer(generalist ~ size + queen + nBees + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
gen16 <- glmer(generalist ~ size + queen  + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
gen17 <- glmer(generalist ~ size  + nBees + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
gen18 <- glmer(generalist ~ ovSavg + queen + nBees + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
gen19 <- glmer(generalist ~ ovSavg + queen  + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
gen20 <- glmer(generalist ~ ovSavg  + nBees + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
gen21 <- glmer(generalist ~ ovRavg + queen + nBees + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
gen22 <- glmer(generalist ~ ovRavg + queen  + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
gen23 <- glmer(generalist ~ ovRavg  + nBees + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
gen24 <- glmer(generalist ~ size + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
gen25 <- glmer(generalist ~ ovSavg + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
gen26 <- glmer(generalist ~ ovRavg + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
gen27 <- glmer(generalist ~ queen + nBees + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
gen28 <- glmer(generalist ~ queen  + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
gen29 <- glmer(generalist ~ nBees + (1|nestID) + (1|natal)+ (1|caste),
              data = count_wide, family = binomial)
model.sel(gen0, gen2, gen5, gen6, gen7, gen8, gen10, gen11, gen13, gen14, gen15, gen16, gen17, gen18, gen19, gen20, gen21, gen22, gen23, gen24, gen25, gen26, gen27, gen28, gen29)

mymodel <- gen24
summary(mymodel)
tab_model(mymodel)
plot(allEffects(mymodel))
```

## lazy
```{r}
# MODEL LAZY
laz0 <- glmer(lazy ~ 1 + (1|natal) + (1|nestID),
              data = count_wide, family = binomial)
laz1 <- glmer(lazy ~ size + ovSavg + ovRavg + queen + nBees + (1|nestID) + (1|natal),
              data = count_wide, family = binomial)
laz2 <- glmer(lazy ~ size + ovSavg + ovRavg + (1|nestID) + (1|natal),
              data = count_wide, family = binomial)
laz3 <- glmer(lazy ~ size + ovSavg + queen + nBees + (1|nestID) + (1|natal),
              data = count_wide, family = binomial)
laz4 <- glmer(lazy ~ size + ovSavg + queen  + (1|nestID) + (1|natal),
              data = count_wide, family = binomial)
laz5 <- glmer(lazy ~ size + ovSavg  + nBees + (1|nestID) + (1|natal),
              data = count_wide, family = binomial)
laz6 <- glmer(lazy ~ size + ovRavg + queen + nBees + (1|nestID) + (1|natal),
              data = count_wide, family = binomial)
laz7 <- glmer(lazy ~ size + ovRavg + queen  + (1|nestID) + (1|natal),
              data = count_wide, family = binomial)
laz8 <- glmer(lazy ~ size + ovRavg  + nBees + (1|nestID) + (1|natal),
              data = count_wide, family = binomial)
laz9 <- glmer(lazy ~ ovSavg + ovRavg + queen + nBees + (1|nestID) + (1|natal),
              data = count_wide, family = binomial)
laz10 <- glmer(lazy ~ ovSavg + ovRavg + queen  + (1|nestID) + (1|natal),
              data = count_wide, family = binomial)
laz11 <- glmer(lazy ~ ovSavg + ovRavg  + nBees + (1|nestID) + (1|natal),
              data = count_wide, family = binomial)
laz12 <- glmer(lazy ~ size + ovSavg + (1|nestID) + (1|natal),
              data = count_wide, family = binomial)
laz13 <- glmer(lazy ~ size + ovRavg + (1|nestID) + (1|natal),
              data = count_wide, family = binomial)
laz14 <- glmer(lazy ~ ovSavg + ovRavg + (1|nestID) + (1|natal),
              data = count_wide, family = binomial)
laz15 <- glmer(lazy ~ size + queen + nBees + (1|nestID) + (1|natal),
              data = count_wide, family = binomial)
laz16 <- glmer(lazy ~ size + queen  + (1|nestID) + (1|natal),
              data = count_wide, family = binomial)
laz17 <- glmer(lazy ~ size  + nBees + (1|nestID) + (1|natal),
              data = count_wide, family = binomial)
laz18 <- glmer(lazy ~ ovSavg + queen + nBees + (1|nestID) + (1|natal),
              data = count_wide, family = binomial)
laz19 <- glmer(lazy ~ ovSavg + queen  + (1|nestID) + (1|natal),
              data = count_wide, family = binomial)
laz20 <- glmer(lazy ~ ovSavg  + nBees + (1|nestID) + (1|natal),
              data = count_wide, family = binomial)
laz21 <- glmer(lazy ~ ovRavg + queen + nBees + (1|nestID) + (1|natal),
              data = count_wide, family = binomial)
laz22 <- glmer(lazy ~ ovRavg + queen  + (1|nestID) + (1|natal),
              data = count_wide, family = binomial)
laz23 <- glmer(lazy ~ ovRavg  + nBees + (1|nestID) + (1|natal),
              data = count_wide, family = binomial)
laz24 <- glmer(lazy ~ size + (1|nestID) + (1|natal),
              data = count_wide, family = binomial)
laz25 <- glmer(lazy ~ ovSavg + (1|nestID) + (1|natal),
              data = count_wide, family = binomial)
laz26 <- glmer(lazy ~ ovRavg + (1|nestID) + (1|natal),
              data = count_wide, family = binomial)
laz27 <- glmer(lazy ~ queen + nBees + (1|nestID) + (1|natal),
              data = count_wide, family = binomial)
laz28 <- glmer(lazy ~ queen  + (1|nestID) + (1|natal),
              data = count_wide, family = binomial)
laz29 <- glmer(lazy ~ nBees + (1|nestID) + (1|natal),
              data = count_wide, family = binomial)
model.sel(laz0, laz2, laz5, laz6, laz7, laz8, laz10, laz11, laz13, laz14, laz15, laz16, laz17, laz18, laz19, laz20, laz21, laz22, laz23, laz24, laz25, laz26, laz27, laz28, laz29)
mymodel <- laz13
summary(mymodel)
tab_model(mymodel)
plot(allEffects(mymodel))
```

## size
```{r}
# MODEL SIZE 
hist(count_data$size)

s0 <- glmer(size ~ 1 + (1|natal) + (1|nestID),
              data = count_data)
s1 <- glmer(size ~ specBehav + ovSavg + ovRavg + queen + nBees + (1|nestID) + (1|natal),
              data = count_data)
s3 <- glmer(size ~ specBehav + ovSavg + ovRavg + (1|nestID) + (1|natal),
              data = count_data)
s4 <- glmer(size ~ specBehav + ovSavg + queen + nBees + (1|nestID) + (1|natal),
              data = count_data)
s5 <- glmer(size ~ specBehav + ovSavg + queen + (1|nestID) + (1|natal),
              data = count_data)
s6 <- glmer(size ~ specBehav + ovSavg + nBees + (1|nestID) + (1|natal),
              data = count_data)
s7 <- glmer(size ~ specBehav + ovRavg + queen + nBees + (1|nestID) + (1|natal),
              data = count_data)
s8 <- glmer(size ~ specBehav + ovRavg + queen + (1|nestID) + (1|natal),
              data = count_data)
s9 <- glmer(size ~ specBehav + ovRavg + nBees + (1|nestID) + (1|natal),
              data = count_data)
s10 <- glmer(size ~ ovSavg + ovRavg + queen + nBees + (1|nestID) + (1|natal),
              data = count_data)
s11 <- glmer(size ~ ovSavg + ovRavg + queen + (1|nestID) + (1|natal),
              data = count_data)
s12 <- glmer(size ~ ovSavg + ovRavg + nBees + (1|nestID) + (1|natal),
              data = count_data)
s13 <- glmer(size ~ specBehav + ovSavg + (1|nestID) + (1|natal),
              data = count_data)
s14 <- glmer(size ~ specBehav + ovRavg + (1|nestID) + (1|natal),
              data = count_data)
s15 <- glmer(size ~ ovSavg + ovRavg + (1|nestID) + (1|natal),
              data = count_data)
s16 <- glmer(size ~ specBehav + queen + nBees + (1|nestID) + (1|natal),
              data = count_data)
s17 <- glmer(size ~ specBehav + queen + (1|nestID) + (1|natal),
              data = count_data)
s18 <- glmer(size ~ specBehav + nBees + (1|nestID) + (1|natal),
              data = count_data)
s19 <- glmer(size ~ ovSavg + queen + nBees + (1|nestID) + (1|natal),
              data = count_data)
s20 <- glmer(size ~ ovSavg + queen + (1|nestID) + (1|natal),
              data = count_data)
s21 <- glmer(size ~ ovSavg + nBees + (1|nestID) + (1|natal),
              data = count_data)
s22 <- glmer(size ~ ovRavg + queen + nBees + (1|nestID) + (1|natal),
              data = count_data)
s23 <- glmer(size ~ ovRavg + queen + (1|nestID) + (1|natal),
              data = count_data)
s24 <- glmer(size ~ ovRavg + nBees + (1|nestID) + (1|natal),
              data = count_data)
s25 <- glmer(size ~ specBehav + (1|nestID) + (1|natal),
              data = count_data)
s26 <- glmer(size ~ ovSavg + (1|nestID) + (1|natal),
              data = count_data)
s27 <- glmer(size ~ ovRavg + (1|nestID) + (1|natal),
              data = count_data)
s28 <- glmer(size ~ queen + nBees + (1|nestID) + (1|natal),
              data = count_data)
s29 <- glmer(size ~ queen + (1|nestID) + (1|natal),
              data = count_data)
s30 <- glmer(size ~ nBees + (1|nestID) + (1|natal),
              data = count_data)
model.sel(s0, s1, s3, s4, s5, s6, s7, s8, s9, s10, s11, s12, s13, s14, s15, s16, s17, s18, s19, s20, s21, s22, s23, s24, s25, s26, s27, s28, s29, s30)
mymodel <- s26
summary(mymodel)
tab_model(mymodel)
plot(allEffects(mymodel))

# plot model residuals
plot(fitted(mymodel), residuals(mymodel), xlab = "Fitted Values", ylab = "Residuals")
abline(h = 0, lty = 2)
lines(smooth.spline(fitted(mymodel), residuals(mymodel)))

# plot all
ggplot(data = count_data, aes(ovRavg, size, color = specBehav, shape = nestTYPE)) +
  geom_point(position = position_jitter(width = 0.1, height = 0.1), size = 2, alpha = 0.7) +
  scale_color_manual(values = c("#F8766D", "#7CAE00", "#00BFC4", "#8c8c8c", "#ffc425")) +
  labs(x = "ovary resorption", y = "body size", color = "behavioral specialization")
ggsave("../figures/ovaryResVsize.jpg")
```

## ovary stage
```{r}
# MODEL OVARY STAGE
hist(count_data$ovSavg)

s0 <- glmer(ovSavg ~ 1 + (1|natal) + (1|nestID),
              data = count_data)
s1 <- glmer(ovSavg ~ specBehav + size + ovRavg + queen + nBees + (1|nestID) + (1|natal),
              data = count_data)
s2 <- glmer(ovSavg ~ specBehav + size + ovRavg + queen + nBees + (1|nestID) + (1|natal),
              data = count_data)
s3 <- glmer(ovSavg ~ specBehav + size + ovRavg + (1|nestID) + (1|natal),
              data = count_data)
s4 <- glmer(ovSavg ~ specBehav + size + queen + nBees + (1|nestID) + (1|natal),
              data = count_data)
s5 <- glmer(ovSavg ~ specBehav + size + queen + (1|nestID) + (1|natal),
              data = count_data)
s6 <- glmer(ovSavg ~ specBehav + size + nBees + (1|nestID) + (1|natal),
              data = count_data)
s7 <- glmer(ovSavg ~ specBehav + ovRavg + queen + nBees + (1|nestID) + (1|natal),
              data = count_data)
s8 <- glmer(ovSavg ~ specBehav + ovRavg + queen + (1|nestID) + (1|natal),
              data = count_data)
s9 <- glmer(ovSavg ~ specBehav + ovRavg + nBees + (1|nestID) + (1|natal),
              data = count_data)
s10 <- glmer(ovSavg ~ size + ovRavg + queen + nBees + (1|nestID) + (1|natal),
              data = count_data)
s11 <- glmer(ovSavg ~ size + ovRavg + queen + (1|nestID) + (1|natal),
              data = count_data)
s12 <- glmer(ovSavg ~ size + ovRavg + nBees + (1|nestID) + (1|natal),
              data = count_data)
s13 <- glmer(ovSavg ~ specBehav + size + (1|nestID) + (1|natal),
              data = count_data)
s14 <- glmer(ovSavg ~ specBehav + ovRavg + (1|nestID) + (1|natal),
              data = count_data)
s15 <- glmer(ovSavg ~ size + ovRavg + (1|nestID) + (1|natal),
              data = count_data)
s16 <- glmer(ovSavg ~ specBehav + queen + nBees + (1|nestID) + (1|natal),
              data = count_data)
s17 <- glmer(ovSavg ~ specBehav + queen + (1|nestID) + (1|natal),
              data = count_data)
s18 <- glmer(ovSavg ~ specBehav + nBees + (1|nestID) + (1|natal),
              data = count_data)
s19 <- glmer(ovSavg ~ size + queen + nBees + (1|nestID) + (1|natal),
              data = count_data)
s20 <- glmer(ovSavg ~ size + queen + (1|nestID) + (1|natal),
              data = count_data)
s21 <- glmer(ovSavg ~ size + nBees + (1|nestID) + (1|natal),
              data = count_data)
s22 <- glmer(ovSavg ~ ovRavg + queen + nBees + (1|nestID) + (1|natal),
              data = count_data)
s23 <- glmer(ovSavg ~ ovRavg + queen + (1|nestID) + (1|natal),
              data = count_data)
s24 <- glmer(ovSavg ~ ovRavg + nBees + (1|nestID) + (1|natal),
              data = count_data)
s25 <- glmer(ovSavg ~ specBehav + (1|nestID) + (1|natal),
              data = count_data)
s26 <- glmer(ovSavg ~ size + (1|nestID) + (1|natal),
              data = count_data)
s27 <- glmer(ovSavg ~ ovRavg + (1|nestID) + (1|natal),
              data = count_data)
s28 <- glmer(ovSavg ~ queen + nBees + (1|nestID) + (1|natal),
              data = count_data)
s29 <- glmer(ovSavg ~ queen + (1|nestID) + (1|natal),
              data = count_data)
s30 <- glmer(ovSavg ~ nBees + (1|nestID) + (1|natal),
              data = count_data)
model.sel(s0, s1, s2, s3, s4, s5, s6, s7, s8, s9, s10, s11, s12, s13, s14, s15, s16, s17, s18, s19, s20, s21, s22, s23, s24, s25, s26, s27, s28, s29, s30)

mymodel <- s15
summary(mymodel)
tab_model(mymodel)
plot(allEffects(mymodel))

# plot model residuals
plot(fitted(mymodel), residuals(mymodel), xlab = "Fitted Values", ylab = "Residuals")
abline(h = 0, lty = 2)
lines(smooth.spline(fitted(mymodel), residuals(mymodel)))



ggplot(data = count_data, aes(ovSavg, size, color = specBehav, shape = nestTYPE)) +
  geom_point(position = position_jitter(width = 0.15, height = 0), size = 2, alpha = 0.7) +
  scale_color_manual(values = c("#F8766D", "#7CAE00", "#00BFC4", "#8c8c8c", "#ffc425")) +
  labs(x = "ovary stage", y = "body size", color = "behavioral specialization")
ggsave("../figures/ovaryVsize.jpg")
```

## ovary resorption
```{r}
# MODEL OVARY RESORPTION
hist(count_data$ovRavg)

s0 <- glmer(ovRavg ~ 1 + (1|natal) + (1|nestID),
              data = count_data)
s1 <- glmer(ovRavg ~ specBehav + size + queen + nBees + (1|nestID) + (1|natal),
              data = count_data)
s2 <- glmer(ovRavg ~ specBehav + size + queen + (1|nestID) + (1|natal),
              data = count_data)
s3 <- glmer(ovRavg ~ specBehav + size + nBees + (1|nestID) + (1|natal),
              data = count_data)
s4 <- glmer(ovRavg ~ specBehav + size + (1|nestID) + (1|natal),
              data = count_data)
s5 <- glmer(ovRavg ~ specBehav + size + queen + nBees + (1|nestID) + (1|natal),
              data = count_data)
s6 <- glmer(ovRavg ~ specBehav + size + queen + (1|nestID) + (1|natal),
              data = count_data)
s7 <- glmer(ovRavg ~ specBehav + size + nBees + (1|nestID) + (1|natal),
              data = count_data)
s8 <- glmer(ovRavg ~ specBehav + queen + nBees + (1|nestID) + (1|natal),
              data = count_data)
s9 <- glmer(ovRavg ~ specBehav + queen + (1|nestID) + (1|natal),
              data = count_data)
s10 <- glmer(ovRavg ~ specBehav + nBees + (1|nestID) + (1|natal),
              data = count_data)
s11 <- glmer(ovRavg ~ size + queen + (1|nestID) + (1|natal),
              data = count_data)
s12 <- glmer(ovRavg ~ specBehav + size + (1|nestID) + (1|natal),
              data = count_data)
s13 <- glmer(ovRavg ~ specBehav + (1|nestID) + (1|natal),
              data = count_data)
s14 <- glmer(ovRavg ~ size + (1|nestID) + (1|natal),
              data = count_data)
s15 <- glmer(ovRavg ~ specBehav + queen + nBees + (1|nestID) + (1|natal),
              data = count_data)
s16 <- glmer(ovRavg ~ specBehav + queen + (1|nestID) + (1|natal),
              data = count_data)
s17 <- glmer(ovRavg ~ specBehav + nBees + (1|nestID) + (1|natal),
              data = count_data)
s18 <- glmer(ovRavg ~ size + queen + nBees + (1|nestID) + (1|natal),
              data = count_data)
s19 <- glmer(ovRavg ~ size + queen + (1|nestID) + (1|natal),
              data = count_data)
s20 <- glmer(ovRavg ~ size + nBees + (1|nestID) + (1|natal),
              data = count_data)
s21 <- glmer(ovRavg ~ queen + nBees + (1|nestID) + (1|natal),
              data = count_data)
s22 <- glmer(ovRavg ~ queen + (1|nestID) + (1|natal),
              data = count_data)
s23 <- glmer(ovRavg ~ nBees + (1|nestID) + (1|natal),
              data = count_data)
s24 <- glmer(ovRavg ~ specBehav + (1|nestID) + (1|natal),
              data = count_data)
s25 <- glmer(ovRavg ~ size + (1|nestID) + (1|natal),
              data = count_data)

model.sel(s0, s1, s2, s3, s4, s5, s6, s7, s8, s9, s10, s11, s12, s13, s14, s15, s16, s17, s18, s19, s20, s21, s22, s23, s24, s25)
mymodel <- s11
summary(mymodel)
tab_model(mymodel)
plot(allEffects(mymodel))
```




# Highest Counts
## Analysis based on highest number of tasks workers perform 
```{r}
head(count_data)
#filter so all bees have >= 3 behaviors (exclude lazies)
high_behav<-count_data %>% filter(lazy=="notLazy") 
nrow(high_behav)
highbehav<-high_behav %>% dplyr::select(beeID, nectarforaging, 
                                 feedsbrood, pollenforaging, egglaying)

hbdf<-highbehav %>% pivot_longer(!beeID, names_to="Behavior", values_to="count")

#Sort each bee by which behavior they do the most and then assign that bee to that behavior
BehavAssign<-hbdf %>% dplyr::group_by(beeID) %>% dplyr::arrange(beeID, desc(count)) %>% dplyr::slice_head() %>% mutate(BehavAssign=Behavior) %>% dplyr::select(beeID, BehavAssign)

#Merge back into original dataframe
FinalBehavRank<-left_join(high_behav, BehavAssign, by="beeID")
FinalBehavRank$BehavAssign <-as.factor(FinalBehavRank$BehavAssign)

#How many bees are in each category? 
nrow(FinalBehavRank[FinalBehavRank$BehavAssign == "nectarforaging",])
nrow(FinalBehavRank[FinalBehavRank$BehavAssign == "pollenforaging",])
nrow(FinalBehavRank[FinalBehavRank$BehavAssign == "feedsbrood",])
nrow(FinalBehavRank[FinalBehavRank$BehavAssign == "egglaying",])

#add highest count to count_data
count_data <- left_join(count_data, BehavAssign, by="beeID")

#plot
ggplot(FinalBehavRank, aes(BehavAssign)) + geom_bar()

p2 <- ggplot(data = count_data, aes(ovSavg, specBehav, color = shannon)) +
  geom_point(position = position_jitter(width=0.1, height=0.2), alpha = 0.8) +
  stat_summary(fun = mean, geom = "point", size = 5) +
  scale_color_gradient(low="dark blue", high="yellow") +
  # scale_color_manual(values = c("#F8766D", "#7CAE00", "#00BFC4", "#8c8c8c", "#ffc425")) +
  labs(x = "ovary stage", y = "behavioral specialization", color = "shannon") 


ggsave("../figures/highcountdistribution.jpg")

#Run GLMMs using BehavAssign as a grouping variable 

B1<-ggplot(FinalBehavRank, aes(BehavAssign, size, color = caste, shape = nestTYPE)) +
  geom_point(position = position_jitter(width=0.1, height=0.05), alpha = 0.5) +
  stat_summary(fun = mean, geom = "point", size = 5) + scale_color_manual(values = c("aquamarine3", "darkslateblue")) + xlab("Most Frequently Performed Task") + 
  ylab("Body Size")
B1
ggsave("../figures/highcountBS.jpg")

FinalBehavRank %>% dplyr::select(BehavAssign, beeID, ovSavg)
B2<-ggplot(FinalBehavRank, aes(BehavAssign, ovSavg, color = caste, shape = nestTYPE)) +
  geom_point(position = position_jitter(width=0.1, height=0.05), alpha = 0.5) +
  stat_summary(fun = mean, geom = "point", size = 5) +scale_color_manual(values = c("aquamarine3", "darkslateblue"))+ xlab("Most Frequently Performed Task") + 
  ylab("Average Ovary \n Development")

B3<-ggplot(FinalBehavRank, aes(BehavAssign, ovRavg, color = caste, shape = nestTYPE)) +
  geom_point(position = position_jitter(width=0.1, height=0.05), alpha = 0.5) +
  stat_summary(fun = mean, geom = "point", size = 5) +scale_color_manual(values = c("aquamarine3", "darkslateblue"))+ xlab("Most Frequently Performed Task") + 
  ylab("Average Ovary Resporption")


ggsave("../figures/highcountOV.jpg")

ggarrange(B1, B2, labels = c("A", "B"), ncol=1, common.legend=TRUE, legend="right") 
ggsave("../figures/highcountsAll.jpg")


```

## Models
```{r}
#Body Size 
hist(FinalBehavRank$size)
head(FinalBehavRank)

s0 <- glmer(size ~ 1 + (1|natal) + (1|nestID) + (1|caste),
              data = FinalBehavRank)
s1 <- glmer(size ~ BehavAssign + (1|nestID) + (1|natal) + (1|caste), data = FinalBehavRank)

model.sel(s0, s1)
# Behavioral group not included in best model 

#Ovary Resporbtion 
hist(FinalBehavRank$ovRavg)
s0 <- glmer(ovRavg ~ 1 + (1|natal) + (1|nestID) + (1|caste),
              data = FinalBehavRank)
s1 <- glmer(ovRavg ~ BehavAssign + (1|nestID) + (1|natal) + (1|caste), data = FinalBehavRank)

model.sel(s0, s1)

#Ovary Stage 
hist(FinalBehavRank$ovSavg)
s0 <- glmer(ovSavg ~ 1 + (1|natal) + (1|nestID) + (1|caste),
              data = FinalBehavRank)
s1 <- glmer(ovSavg ~ BehavAssign + (1|nestID) + (1|natal) + (1|caste), data = FinalBehavRank)

model.sel(s0, s1)

#Behaviorlal assign not included in any of the best models 
```



# Perfectly specialized counts- size and ovaries 
```{r}
# subset to look only at perfectly specialized individuals
perfSpec <- count_data %>% filter(lazy=="notLazy") %>% filter(shannon=="0")
# How many bees perform more than one task? 
1-nrow(perfSpec)/nrow(count_data)

nrow(perfSpec[perfSpec$specBehav == "nectarforaging",])
nrow(perfSpec[perfSpec$specBehav == "pollenforaging",])
nrow(perfSpec[perfSpec$specBehav == "feedsbrood",])
nrow(perfSpec[perfSpec$specBehav == "egglaying",])

perfgraph<-perfSpec %>% dplyr::select(specBehav, beeID, ovSavg)
perfgraph <-perfgraph[complete.cases(perfgraph),]

perfSpec

# add perfSpec to count_data
count_data$perfSpec <- NA
for (i in 1:nrow(count_data)) {
  if (is.na(count_data$specProp[i])) {
  } else if (count_data$specProp[i] == 1) {
    count_data$perfSpec[i] <- count_data$specBehav[i] 
    }
}
count_data
# write count_data for nmds graphs
write.csv(count_data, "../Data/filtData/count_data.csv")

graphperfSpec<-perfgraph %>% group_by(specBehav) %>% 
  mutate(aov=mean(ovSavg))  %>% 
  mutate(sdov=sd(ovSavg)) %>% 
  dplyr::select(specBehav, aov, sdov) %>% unique()
graphperfSpec

perfov<-ggplot(graphperfSpec, aes(x=specBehav, y=aov)) +
    geom_bar(stat="identity", position=position_dodge(), colour="black") + 
  geom_errorbar(aes(ymin=aov-sdov,  ymax=aov+sdov))
perfov

```


## Models 
```{r}
#Body Size 
hist(perfSpec$size)

s0 <- glmer(size ~ 1 + (1|natal) + (1|nestID) + (1|caste),
              data = perfSpec)
s1 <- glmer(size ~ specBehav + (1|nestID) + (1|natal) + (1|caste), data = perfSpec)

model.sel(s0, s1)

#Ovary Resporbtion 
hist(perfSpec$ovRavg)
s0 <- glmer(ovRavg ~ 1 + (1|natal) + (1|nestID) + (1|caste),
              data = perfSpec)
s1 <- glmer(ovRavg ~ specBehav + (1|nestID) + (1|natal) + (1|caste), data = perfSpec)

model.sel(s0, s1)

#Ovary Stage 
hist(FinalBehavRank$ovSavg)
s0 <- glmer(ovSavg ~ 1 + (1|natal) + (1|nestID) + (1|caste),
              data = perfSpec)
s1 <- glmer(ovSavg ~ specBehav + (1|nestID) + (1|natal) + (1|caste), data = perfSpec)

model.sel(s0, s1)

mymodel <- s1
summary(mymodel)
tab_model(mymodel)
plot(allEffects(mymodel))

# plot model residuals
plot(fitted(mymodel), residuals(mymodel), xlab = "Fitted Values", ylab = "Residuals")
abline(h = 0, lty = 2)
lines(smooth.spline(fitted(mymodel), residuals(mymodel)))

posthoc<-glht(s1, linfct = mcp(specBehav = "Tukey"))
summary(posthoc)
perfSpec$nestTYPE=as.factor(perfSpec$nestTYPE)
perfSpec$caste=as.factor(perfSpec$caste)

C1<-ggplot(data = perfSpec, aes(specBehav, ovSavg, color = caste, shape=nestTYPE)) +
  geom_point(position = position_jitter(width = 0.1, height = 0.1), size = 2, alpha = 0.7) +labs(x = "Perfectly Specialized Behavior", y = "Average Ovary \n Stage", shape = "Nest Configuration")+ stat_summary(fun = mean, geom = "point", size = 5)+ scale_color_manual(values = c("aquamarine3", "darkslateblue"))
ggsave("../figures/osperfectlyspecialized.jpg")


```






## Body Size 
```{r}
# MODEL SIZE 
hist(perfSpec$size)

s0 <- glmer(size ~ 1 + (1|natal) + (1|nestID),
              data = perfSpec)
s1 <- glmer(size ~ specBehav + ovSavg + ovRavg + queen + nBees + (1|nestID) + (1|natal),
              data = perfSpec)
s3 <- glmer(size ~ specBehav + ovSavg + ovRavg + (1|nestID) + (1|natal),
              data = perfSpec)
s4 <- glmer(size ~ specBehav + ovSavg + queen + nBees + (1|nestID) + (1|natal),
              data = perfSpec)
s5 <- glmer(size ~ specBehav + ovSavg + queen + (1|nestID) + (1|natal),
              data = perfSpec)
s6 <- glmer(size ~ specBehav + ovSavg + nBees + (1|nestID) + (1|natal),
              data = perfSpec)
s7 <- glmer(size ~ specBehav + ovRavg + queen + nBees + (1|nestID) + (1|natal),
              data = perfSpec)
s8 <- glmer(size ~ specBehav + ovRavg + queen + (1|nestID) + (1|natal),
              data = perfSpec)
s9 <- glmer(size ~ specBehav + ovRavg + nBees + (1|nestID) + (1|natal),
              data = perfSpec)
s10 <- glmer(size ~ ovSavg + ovRavg + queen + nBees + (1|nestID) + (1|natal),
              data = perfSpec)
s11 <- glmer(size ~ ovSavg + ovRavg + queen + (1|nestID) + (1|natal),
              data = perfSpec)
s12 <- glmer(size ~ ovSavg + ovRavg + nBees + (1|nestID) + (1|natal),
              data = perfSpec)
s13 <- glmer(size ~ specBehav + ovSavg + (1|nestID) + (1|natal),
              data = perfSpec)
s14 <- glmer(size ~ specBehav + ovRavg + (1|nestID) + (1|natal),
              data = perfSpec)
s15 <- glmer(size ~ ovSavg + ovRavg + (1|nestID) + (1|natal),
              data = perfSpec)
s16 <- glmer(size ~ specBehav + queen + nBees + (1|nestID) + (1|natal),
              data = perfSpec)
s17 <- glmer(size ~ specBehav + queen + (1|nestID) + (1|natal),
              data = perfSpec)
s18 <- glmer(size ~ specBehav + nBees + (1|nestID) + (1|natal),
              data = perfSpec)
s19 <- glmer(size ~ ovSavg + queen + nBees + (1|nestID) + (1|natal),
              data = perfSpec)
s20 <- glmer(size ~ ovSavg + queen + (1|nestID) + (1|natal),
              data = perfSpec)
s21 <- glmer(size ~ ovSavg + nBees + (1|nestID) + (1|natal),
              data = perfSpec)
s22 <- glmer(size ~ ovRavg + queen + nBees + (1|nestID) + (1|natal),
              data = perfSpec)
s23 <- glmer(size ~ ovRavg + queen + (1|nestID) + (1|natal),
              data = perfSpec)
s24 <- glmer(size ~ ovRavg + nBees + (1|nestID) + (1|natal),
              data = perfSpec)
s25 <- glmer(size ~ specBehav + (1|nestID) + (1|natal),
              data = perfSpec)
s26 <- glmer(size ~ ovSavg + (1|nestID) + (1|natal),
              data = perfSpec)
s27 <- glmer(size ~ ovRavg + (1|nestID) + (1|natal),
              data = perfSpec)
s28 <- glmer(size ~ queen + nBees + (1|nestID) + (1|natal),
              data = perfSpec)
s29 <- glmer(size ~ queen + (1|nestID) + (1|natal),
              data = perfSpec)
s30 <- glmer(size ~ nBees + (1|nestID) + (1|natal),
              data = perfSpec)
model.sel(s0, s1, s2, s3, s4, s5, s6, s7, s8, s9, s10, s11, s12, s13, s14, s15, s16, s17, s18, s19, s20, s21, s22, s23, s24, s25, s26, s27, s28, s29, s30)

s01 <- lmer(size ~ specBehav + (1|nestTYPE) + (1|nestID) + (1|natal) + (1|caste),
              data = perfSpec)

mymodel <- s01
summary(mymodel)
tab_model(mymodel)
plot(allEffects(mymodel))

# plot model residuals
plot(fitted(mymodel), residuals(mymodel), xlab = "Fitted Values", ylab = "Residuals")
abline(h = 0, lty = 2)
lines(smooth.spline(fitted(mymodel), residuals(mymodel)))

posthoc<-glht(s01, linfct = mcp(specBehav = "Tukey"))
summary(posthoc)


# plot size against behavioral specialization 

C2<-ggplot(data = perfSpec, aes(specBehav, size, color = caste, shape=nestTYPE)) +
  geom_point(position = position_jitter(width = 0.1, height = 0.1), size = 2, alpha = 0.7) +labs(x = "Perfectly Specialized Behavior", y = "Body Size", shape = "Nest Configuration")+ stat_summary(fun = mean, geom = "point", size = 5)+ scale_color_manual(values = c("aquamarine3", "darkslateblue"))
ggsave("../figures/bsperfectlyspecialized.jpg")

ggarrange(C2, C1, labels = c("A", "B"), ncol=1, common.legend=TRUE, legend="right")
ggsave("../figures/perfectcountsAll.jpg")


```


```{r}
# plot all
ggplot(data = perfSpec, aes(ovRavg, size, color = specBehav, shape = nestTYPE)) +
  geom_point(position = position_jitter(width = 0.1, height = 0.1), size = 2, alpha = 0.7) +
  scale_color_manual(values = c("#F8766D", "#7CAE00", "#00BFC4", "#8c8c8c", "#ffc425")) +
  labs(x = "ovary resorption", y = "body size", color = "behavioral specialization")
ggsave("../figures/ovaryResVsize.jpg")





# MODEL OVARY STAGE
hist(perfSpec$ovSavg)

s0 <- glmer(ovSavg ~ 1 + (1|natal) + (1|nestID),
              data = perfSpec)
s1 <- glmer(ovSavg ~ specBehav + size + ovRavg + queen + nBees + (1|nestID) + (1|natal),
              data = perfSpec)
s2 <- glmer(ovSavg ~ specBehav + size + ovRavg + queen + nBees + (1|nestID) + (1|natal),
              data = perfSpec)
s3 <- glmer(ovSavg ~ specBehav + size + ovRavg + (1|nestID) + (1|natal),
              data = perfSpec)
s4 <- glmer(ovSavg ~ specBehav + size + queen + nBees + (1|nestID) + (1|natal),
              data = perfSpec)
s5 <- glmer(ovSavg ~ specBehav + size + queen + (1|nestID) + (1|natal),
              data = perfSpec)
s6 <- glmer(ovSavg ~ specBehav + size + nBees + (1|nestID) + (1|natal),
              data = perfSpec)
s7 <- glmer(ovSavg ~ specBehav + ovRavg + queen + nBees + (1|nestID) + (1|natal),
              data = perfSpec)
s8 <- glmer(ovSavg ~ specBehav + ovRavg + queen + (1|nestID) + (1|natal),
              data = perfSpec)
s9 <- glmer(ovSavg ~ specBehav + ovRavg + nBees + (1|nestID) + (1|natal),
              data = perfSpec)
s10 <- glmer(ovSavg ~ size + ovRavg + queen + nBees + (1|nestID) + (1|natal),
              data = perfSpec)
s11 <- glmer(ovSavg ~ size + ovRavg + queen + (1|nestID) + (1|natal),
              data = perfSpec)
s12 <- glmer(ovSavg ~ size + ovRavg + nBees + (1|nestID) + (1|natal),
              data = perfSpec)
s13 <- glmer(ovSavg ~ specBehav + size + (1|nestID) + (1|natal),
              data = perfSpec)
s14 <- glmer(ovSavg ~ specBehav + ovRavg + (1|nestID) + (1|natal),
              data = perfSpec)
s15 <- glmer(ovSavg ~ size + ovRavg + (1|nestID) + (1|natal),
              data = perfSpec)
s16 <- glmer(ovSavg ~ specBehav + queen + nBees + (1|nestID) + (1|natal),
              data = perfSpec)
s17 <- glmer(ovSavg ~ specBehav + queen + (1|nestID) + (1|natal),
              data = perfSpec)
s18 <- glmer(ovSavg ~ specBehav + nBees + (1|nestID) + (1|natal),
              data = perfSpec)
s19 <- glmer(ovSavg ~ size + queen + nBees + (1|nestID) + (1|natal),
              data = perfSpec)
s20 <- glmer(ovSavg ~ size + queen + (1|nestID) + (1|natal),
              data = perfSpec)
s21 <- glmer(ovSavg ~ size + nBees + (1|nestID) + (1|natal),
              data = perfSpec)
s22 <- glmer(ovSavg ~ ovRavg + queen + nBees + (1|nestID) + (1|natal),
              data = perfSpec)
s23 <- glmer(ovSavg ~ ovRavg + queen + (1|nestID) + (1|natal),
              data = perfSpec)
s24 <- glmer(ovSavg ~ ovRavg + nBees + (1|nestID) + (1|natal),
              data = perfSpec)
s25 <- glmer(ovSavg ~ specBehav + (1|nestID) + (1|natal),
              data = perfSpec)
s26 <- glmer(ovSavg ~ size + (1|nestID) + (1|natal),
              data = perfSpec)
s27 <- glmer(ovSavg ~ ovRavg + (1|nestID) + (1|natal),
              data = perfSpec)
s28 <- glmer(ovSavg ~ queen + nBees + (1|nestID) + (1|natal),
              data = perfSpec)
s29 <- glmer(ovSavg ~ queen + (1|nestID) + (1|natal),
              data = perfSpec)
s30 <- glmer(ovSavg ~ nBees + (1|nestID) + (1|natal),
              data = perfSpec)
model.sel(s0, s1, s2, s3, s4, s5, s6, s7, s8, s9, s10, s11, s12, s13, s14, s15, s16, s17, s18, s19, s20, s21, s22, s23, s24, s25, s26, s27, s28, s29, s30)


st01 <- glmer(ovSavg ~ specBehav + (1|nestTYPE) + (1|nestID) + (1|natal),
              data = perfSpec, family = gaussian(link="identity"))


mymodel <- st01
summary(mymodel)
tab_model(mymodel)
plot(allEffects(mymodel))

posthoc <- glht(mymodel, linfct = mcp(specBehav = "Tukey"))
posthoc

# plot model residuals
plot(fitted(mymodel), residuals(mymodel), xlab = "Fitted Values", ylab = "Residuals")
abline(h = 0, lty = 2)
lines(smooth.spline(fitted(mymodel), residuals(mymodel)))

ggplot(data = perfSpec, aes(ovSavg, size, color = specBehav, shape = nestTYPE)) +
  geom_point(position = position_jitter(width = 0.15, height = 0), size = 2, alpha = 0.7) +
  scale_color_manual(values = c("#F8766D", "#7CAE00", "#00BFC4", "#8c8c8c", "#ffc425")) +
  labs(x = "ovary stage", y = "body size", color = "behavioral specialization")
ggsave("../figures/ovaryVsize.jpg")




# MODEL OVARY RESORPTION
hist(perfSpec$ovRavg)

s0 <- glmer(ovRavg ~ 1 + (1|natal) + (1|nestID),
              data = perfSpec)
s1 <- glmer(ovRavg ~ specBehav + size + queen + nBees + (1|nestID) + (1|natal),
              data = perfSpec)
s2 <- glmer(ovRavg ~ specBehav + size + queen + (1|nestID) + (1|natal),
              data = perfSpec)
s3 <- glmer(ovRavg ~ specBehav + size + nBees + (1|nestID) + (1|natal),
              data = perfSpec)
s4 <- glmer(ovRavg ~ specBehav + size + (1|nestID) + (1|natal),
              data = perfSpec)
s5 <- glmer(ovRavg ~ specBehav + size + queen + nBees + (1|nestID) + (1|natal),
              data = perfSpec)
s6 <- glmer(ovRavg ~ specBehav + size + queen + (1|nestID) + (1|natal),
              data = perfSpec)
s7 <- glmer(ovRavg ~ specBehav + size + nBees + (1|nestID) + (1|natal),
              data = perfSpec)
s8 <- glmer(ovRavg ~ specBehav + queen + nBees + (1|nestID) + (1|natal),
              data = perfSpec)
s9 <- glmer(ovRavg ~ specBehav + queen + (1|nestID) + (1|natal),
              data = perfSpec)
s10 <- glmer(ovRavg ~ specBehav + nBees + (1|nestID) + (1|natal),
              data = perfSpec)
s11 <- glmer(ovRavg ~ size + queen + (1|nestID) + (1|natal),
              data = perfSpec)
s12 <- glmer(ovRavg ~ specBehav + size + (1|nestID) + (1|natal),
              data = perfSpec)
s13 <- glmer(ovRavg ~ specBehav + (1|nestID) + (1|natal),
              data = perfSpec)
s14 <- glmer(ovRavg ~ size + (1|nestID) + (1|natal),
              data = perfSpec)
s15 <- glmer(ovRavg ~ specBehav + queen + nBees + (1|nestID) + (1|natal),
              data = perfSpec)
s16 <- glmer(ovRavg ~ specBehav + queen + (1|nestID) + (1|natal),
              data = perfSpec)
s17 <- glmer(ovRavg ~ specBehav + nBees + (1|nestID) + (1|natal),
              data = perfSpec)
s18 <- glmer(ovRavg ~ size + queen + nBees + (1|nestID) + (1|natal),
              data = perfSpec)
s19 <- glmer(ovRavg ~ size + queen + (1|nestID) + (1|natal),
              data = perfSpec)
s20 <- glmer(ovRavg ~ size + nBees + (1|nestID) + (1|natal),
              data = perfSpec)
s21 <- glmer(ovRavg ~ queen + nBees + (1|nestID) + (1|natal),
              data = perfSpec)
s22 <- glmer(ovRavg ~ queen + (1|nestID) + (1|natal),
              data = perfSpec)
s23 <- glmer(ovRavg ~ nBees + (1|nestID) + (1|natal),
              data = perfSpec)
s24 <- glmer(ovRavg ~ specBehav + (1|nestID) + (1|natal),
              data = perfSpec)
s25 <- glmer(ovRavg ~ size + (1|nestID) + (1|natal),
              data = perfSpec)

model.sel(s0, s1, s2, s3, s4, s5, s6, s7, s8, s9, s10, s11, s12, s13, s14, s15, s16, s17, s18, s19, s20, s21, s22, s23, s24, s25)

r01 <- glmer(ovRavg ~ specBehav + (1|nestTYPE) + (1|nestID) + (1|natal),
              data = perfSpec)

mymodel <- r01
summary(mymodel)
tab_model(mymodel)
plot(allEffects(mymodel))
```

```
