---
title: "nmdsAll"
author: "ES"
date: "7/30/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# load packages
```{r}
library(tidyverse)
library(dplyr)
library(plyr)
library(ggplot2)
library(cluster) # clustering algorithms
library(factoextra)
library(vegan) 
library(lme4)
library(plotly)
```

* you must run the 4_shannonBees.Rmd file before running this. 

# load data
```{r}
# use bee counts filt (bees with >=3 obs each pooled across time)
nmds_data <- read.csv("../Data/tempData/beeCounts_filt.csv")

# add specialist information from shannonBees.Rmd file
spec <- read.csv("../Data/tempData/beeCounts_filtSpecPOOLED.csv")

# add ovary and body size information to analyze specialists 
ovary <- read.csv("../Data/Ovarydata.csv", col.names = c("nestID", "beeID", "fb", "ov", "l1", "s1", "l2", "s2", "notes", "x", "xx"))
size <- read.csv("../Data/bodysizeworkers.csv", col.names = c("beeID", "nestID", "date", "wing1", "wing2", "size"), na.strings = "cell missing")
# make beeID unique
ovary$beeID <- paste0(ovary$nestID, ovary$beeID)
size$beeID <- paste0(size$nestID, size$beeID)
ovary
```

# NMDS on all pooled data -- individual specialization over time
```{r}
# make NA values 0
nmds_data[is.na(nmds_data)] <- 0
nmds_data
numBees <- nrow(nmds_data)
# remove bees lacking any behaviors
nmds_data <-nmds_data[rowSums(nmds_data[(10:(ncol(nmds_data) - 1))]) > 0, ]
numDatapoints <- nrow(nmds_data)
print(paste( (numBees - numDatapoints), "bees were removed from the nmds data because they had 0 recorded behaviors", sep = " "))

# # add dummy variable to include lazy workers
# nmds_data$nothing <- 0.01

# remove bees with NA beeID
nmds_data <- nmds_data[!is.na(nmds_data$beeID),]

# make matrix of only behavior counts for all timepoints
nmds_all <- as.matrix(nmds_data[, c("feedsbrood", "nectarforaging", "pollenforaging", "egglaying")]) 
nmds_all

# Run NMDS with euclidean distance
nmds_all <- metaMDS(nmds_all)
plot(nmds_all)

# Get coordinates to graph it 
# Gets NMDS info for each individual
data.scores <- as.data.frame(scores(nmds_all))
data.scores$caste <- nmds_data$caste
data.scores$nestTYPE <-nmds_data$nestTYPE
data.scores$beeID <-nmds_data$beeID
data.scores$nestID <- nmds_data$nestID
data.scores$nBees <- nmds_data$nBees
data.scores$nWork <- nmds_data$nWork
data.scores$queen <- nmds_data$queen
data.scores

# Gets NMDS info for each behavior
behav.scores<-as.data.frame(scores(nmds_all, "species"))
behav.scores$behav<-rownames(behav.scores)
behav.scores

data.scores$nestTYPE <- ordered(data.scores$nestTYPE, levels = c("NT", "QT", "NF", "QF"))

# Plot it using ggplot
ggplot() + 
  geom_point(data=data.scores, aes(x=NMDS1, y=NMDS2, shape = caste, color = nestTYPE), size = 2, position = position_jitter(width=0.05, height=0.05), alpha = 0.6) + 
  scale_shape_manual(values = c(21, 16)) +
  geom_text(data=behav.scores, aes(x=NMDS1, y=NMDS2, label=behav)) +
  coord_cartesian(xlim=c(-1.5, 1.5), ylim=c(-1.5, 1.5))
ggsave("../figures/nmdsAll.jpg")
```

# add specialists onto nmds plot
***NEED TO RUN SHANNONBEES RMD FILE FIRST TO ACQUIRE BEECOUNTS_FILTSPECPOOLED.CSV
```{r}
# make beeID include nestID so we can identify individually
spec$beeID <- paste0(spec$nestID, spec$beeID)
data.scores$beeID <- paste0(data.scores$nestID, data.scores$beeID)

# make NA specialists named "none"
spec[is.na(spec$specBehav), "specBehav"] <- "none"

# make new column for specBehav in data.scores
data.scores$specBehav <- NA
# go through each bee we have an nmds score for and add specialist category if applicable
for (i in 1:nrow(data.scores)) {
  for (j in 1:nrow(spec)) {
    if (data.scores$beeID[i] == spec$beeID[j]) {
      data.scores$specBehav[i] <- spec$specBehav[j]
    }
  }
}
data.scores
data.scores[data.scores$specBehav == "nectarforaging", ]

data.scores$specBehav <- ordered(data.scores$specBehav, levels = c("egglaying", "feedsbrood", "nectarforaging", "pollenforaging", "none"))
behav.scores$behav <- ordered(behav.scores$behav, levels = c("egglaying", "feedsbrood", "nectarforaging", "pollenforaging", "none"))

# Plot it using ggplot
ggplot() + 
  geom_point(data=data.scores, aes(x=NMDS1, y=NMDS2, shape = caste, color = specBehav), size = 2, position = position_jitter(width=0.05, height=0.05), alpha = 0.7) + 
  scale_shape_manual(values = c(21, 16)) +
  scale_color_manual(values = c("#C77CFF", "#00BFC4", "#F8766D", "grey")) +
  geom_text(data=behav.scores, aes(x=NMDS1, y=NMDS2, label=behav)) +
  coord_cartesian(xlim=c(-1.5, 1.5), ylim=c(-1.5, 2)) +
  labs(color = "specialist")
ggsave("../figures/nmdsSpec.jpg")
```

# does body size correspond to behavioral specialization?
```{r}
size
data.scores

# add body size to data.scores df
for (i in 1:nrow(data.scores)) {
  for (j in 1:nrow(size)) {
    if (data.scores$beeID[i] == size$beeID[j]) {
      data.scores$size[i] <- size$size[j]
    }
  }
}

# add workers without behaviors
# subset size to look only at bees missing from data.scores (these are bees with <3 obs)
lazy <-  subset(size, !(beeID %in% data.scores$beeID))
# add that number of rows to data.scores
addLazy <- matrix(NA, nrow = nrow(lazy), ncol = ncol(data.scores))
addLazy <- as.data.frame(addLazy)
colnames(addLazy) <- colnames(data.scores)
# add beeID, nestID and size of lazy bees to addLazy
addLazy$beeID <- lazy$beeID
addLazy$nestID <- lazy$nestID
addLazy$size <- lazy$size
addLazy$specBehav <- "lazy"
# bind to data.scores
data.scores <- rbind(data.scores, addLazy)

addLazy
lazy
data.scores
ggplot(data = data.scores, aes(size, specBehav, color = specBehav)) +
  geom_point(position = position_jitter(width=0.1, height=0.2), alpha = 0.4) +
  stat_summary(fun = mean, geom = "point", size = 5) +
  labs(x = "body size", y = "behavioral specialization", color = "behavioral specialization")
ggsave("../figures/bodySize.jpg")
```

# does ovary development correspond to behavioral specialization?
```{r}
ovary
data.scores

# calculate average ovary length
ovary$length <- rowMeans(ovary[,c("l1", "l2")], na.rm = TRUE)

# add ovary data to data.scores df
for (i in 1:nrow(data.scores)) {
  for (j in 1:nrow(ovary)) {
    if (data.scores$beeID[i] == ovary$beeID[j]) {
      data.scores$length[i] <- ovary$length[j]
    }
  }
}

data.scores

ggplot(data = data.scores, aes(length, specBehav, color = specBehav)) +
  geom_point(position = position_jitter(width=0.1, height=0.2), alpha = 0.4) +
  stat_summary(fun = mean, geom = "point", size = 5) +
  labs(x = "ovary length", y = "behavioral specialization", color = "behavioral specialization")
ggsave("../figures/ovaryLength.jpg")
```












