---
title: "CleanData"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
  
# Load Packages
```{r}
library(tidyverse)
```

# Load Data
```{r}

## create a list of all the files in the BORISfiles directory. these are all the files we want to import
## directory will change depending on user, to find your file path, use command "pwd" 
borisfiles <- list.files("/Users/kaleighfisher/Google Drive File Stream/My Drive/WoodardLab.shared/Projects/2018.Awesome.Summer.Experiments/Raw Data") 


## loop over the files and import all of them. rename column names to remove "." and make shorter
lapply(borisfiles, read.csv, header = TRUE, col.names = c("obsID", "obsDate", "filename", "totalLen", "fps", "beeID", "behav", "mod", "behavType", "start", "stop", "dur", "comStart", "comStop"))

## merge the files into one dataframe (do.call does stuff to all the things)
## this is NOT CORRECT! need to edit to make it work
do.call(rbind(filename))




##
## load a subset of data to test
all_data <- read.csv("/Users/ericasarro/Google\ Drive\ File\ Stream/My\ Drive/MasterProjects/18Su_AwesomeSummer/MockData/AggEvents.csv", header = TRUE, col.names = c("obsID", "obsDate", "filename", "totalLen", "fps", "beeID", "behav", "mod", "behavType", "start", "stop", "dur", "comStart", "comStop" ))
head(all_data)

```

# Clean Data
```{r}

## convert columns to character to enable manipulation with strsplit() and subset(startsWith())
all_data$beeID <- as.character(all_data$beeID)
all_data$obsID <- as.character(all_data$obsID)

## split out obsID into camID, nestID, dateID and hourID (split based on the tab (\t) between elements)
## element 1 is camID
all_data$camID <- sapply(strsplit(all_data$obsID, "\t"), 
                          function(x){
                            x[1]
                            } )
## element 2 is nestID
all_data$nestID <- sapply(strsplit(all_data$obsID, "\t"), 
                          function(x){
                            x[2]
                            } )
## element 3 is dateID
all_data$dateID <- sapply(strsplit(all_data$obsID, "\t"), 
                          function(x){ 
                            x[3]
                            } )
## element 4 is hourID
all_data$hourID <- sapply(strsplit(all_data$obsID, "\t"), 
                          function(x){ 
                            x[4]
                            } )
```

```{r}
## collapse the dataframe to count the number of instances doing each behavior
## this outputs a new dataframe with columns nestID, beeID, nectarforaging, feedsbrood, pollenforaging 
#  (and any other behaviors recorded), with counts for each behavior in as values
## produces error: Values in `dur` are not uniquely identified; output will contain list-cols. 
new_data <- all_data %>%
  pivot_wider(id_cols = c("nestID", "beeID"), names_from = "behav", values_from = "dur") 
new_data


```

```{r}
## group data by subject
subj_data <- group_by(all_data, beeID)

## group data by behavior
behav_data <- group_by(subj_data, behav)

## calculate absolute amount of time spent on each behav
#  for each subject: sum time carrying out each task, divide by total time watched for that nestID
#  when calculating total time:
#   **need to account for videos with 0 data. make sure boris output includes those
#   **need to account for multiple events in a single video -- can only sum unique obsIDs

## for each beeID, count the number of unique behaviors
sapply(behav_data$beeID, sapply(behav_data$behav, count(behav_data$behav)))

for (i in levels(behav_data$beeID)) {
  behav_beeID <- subset(behav_data, beeID = i)
  num_feeds <- sum(behav_beeID$behav == "feedsbrood")
  return(num_feeds)
}

num_feeds


       for (all unique subjects) {
  for (all unique behaviors) {
    sum number of behav
    sum time of dur
  }
}
    
