---
title: "filtering"
author: "ES"
date: "6/22/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# load packages
```{r}
library(ggplot2)
library(tidyr)
library(dplyr)
```

# load data
```{r}
counts_data <- read.csv("../Data/countsanalysis.csv")
counts_data
```

# look at the number of events of each behavior per nest
# filter out nests with < 3 observations each of brood feeding and foraging behaviors
counts_filt = nests with at least 3 observations of each behavior
*can be used in all analyses in which timepoints are pooled
```{r}
# first combine pollen and nectar foraging into a single value
counts_data$allforaging <- counts_data$nectarforaging + counts_data$pollenforaging

# gather all behaviors into a single column
counts_gathered <- counts_data %>% gather("feedsbrood", "allforaging", key = "behav", value = "count")
counts_gathered

# sumNest <- counts_gathered %>% group_by(nestID, behav)
# 
# ggplot(sumNest, aes(x = nestID, y = count, fill = behav)) +
#   geom_bar(stat ="identity") + 
#   theme(axis.text.x = element_text(angle = 90), legend.position = "top")

# summarize the total behaviors observed across all individuals
statsBehav <- ddply(counts_gathered, 
                c("behav"), 
                summarise,
                nObs = sum(count),
                mean = mean(count), 
                sd = sd(count),
                sem = sd(count)/sqrt(length(count)))
statsBehav

PstatsBehav <- ggplot(statsBehav, aes(x = behav, y = nObs, fill = behav)) + geom_bar(stat = "identity")
PstatsBehav

# summarize behaviors observed for each nest
statsNest <- ddply(counts_gathered, 
                c("nestID", "behav"), 
                summarise,
                nObs = sum(count),
                mean = mean(count), 
                sd = sd(count),
                sem = sd(count)/sqrt(length(count)))
statsNest
PstatsNest <- ggplot(statsNest, aes(x = nestID, y = nObs, fill = behav)) + 
  geom_bar(stat = "identity") + 
  theme(axis.text.x = element_text(angle = 90), legend.position = "top")
PstatsNest

# remove nests with fewer than 3 observations each of foraging and brood feeding
# look at which nests need to be removed
rmObs <- statsNest[statsNest$nObs < 3,]
rmObs
rmNests <- unique(rmObs$nestID)
print(paste(length(rmNests), "nests need to be removed from the pooled time data because they have fewer than 3 instances of brood feeding or foraging recorded"))
rmNests
counts_filt <- counts_data %>% filter(!nestID %in% rmNests)
counts_filt
write.csv(counts_filt, "../Data/tempData/counts_filt.csv")
```

# do the same for early and late timepoints individually
counts_time = count data with at least 3 observations of each behavior at both early and late timepoints
*can be used in all analyses that compare across time
```{r}
# summarize behaviors observed for each nest at early and late timepoints
statsNestTime <- ddply(counts_gathered, 
                c("nestID", "behav", "timepoint"), 
                summarise,
                nObs = sum(count),
                mean = mean(count), 
                sd = sd(count),
                sem = sd(count)/sqrt(length(count)))

# summarize early timepoint
statsNestEarly <- statsNestTime[statsNestTime$timepoint == "early",]
PstatsNestEarly <- ggplot(statsNestEarly, aes(x = nestID, y = nObs, fill = behav)) + 
  geom_bar(stat = "identity") + 
  theme(axis.text.x = element_text(angle = 90), legend.position = "top")
PstatsNestEarly
# remove nests wither fewer than 3 observations per behavior at early timepoint
rmObsEarly <- statsNestEarly[statsNestEarly$nObs < 3,]
rmObsEarly
rmNestsEarly <- unique(rmObsEarly$nestID)
print(paste(length(rmNestsEarly), "nests need to be removed from the early timepoint data because they have fewer than 3 instances of brood feeding or foraging recorded"))
rmNestsEarly
goodNestsEarly <- counts_data %>% filter(!nestID %in% rmNestsEarly)
unique(goodNestsEarly$nestID)

# summarize late timepoint
statsNestLate <- statsNestTime[statsNestTime$timepoint == "late",]
PstatsNestLate <- ggplot(statsNestLate, aes(x = nestID, y = nObs, fill = behav)) + 
  geom_bar(stat = "identity") + 
  theme(axis.text.x = element_text(angle = 90), legend.position = "top")
PstatsNestLate
# remove nests wither fewer than 3 observations per behavior at late timepoint
rmObsLate <- statsNestLate[statsNestLate$nObs < 3,]
rmObsLate
rmNestsLate <- unique(rmObsLate$nestID)
print(paste(length(rmNestsLate), "nests need to be removed from the late timepoint data because they have fewer than 3 instances of brood feeding or foraging recorded"))
rmNestsLate
goodNestsLate <- counts_data %>% filter(!nestID %in% rmNestsLate)
goodNestsLate
unique(goodNestsLate$nestID)

# dataframe that can be used to compare across timepoints (at least 3 observations of each behavior in both early and late timepoints)
counts_time <- counts_data %>% filter(!nestID %in% rmNestsEarly) %>% filter(!nestID %in% rmNestsLate)
counts_time
unique(counts_time$nestID)
write.csv(counts_time, "../Data/tempData/counts_time.csv")
```