---
title: "Bumble Bee Queen Time Budgeting and Task Repertoire"
output: html_notebook
---

# Proposed Topic
How does the social environment affect bumble bee queen task repertoire and time budget?


# Background
Bumble bees have an annual life cycle
-contains both a solitary phase and a social phase

Solitary Phase: queens exhibit broad task repertoire, carry out brood care, nest maintenance, and foraging tasks independently

Social Phase: queens transition to focusing almost exclusively on reproduction. workers take over brood care and foraging tasks


# Data Collection Methods
-Q0 queens (solitary)
-Q3 queens (with 3 workers)
-Q5 queens (with 5 workers)

-2 time points per colony (early + late)

-score behavioral data in videos


# Video Data
Data includes several spreadsheets (each spreadsheet corresponding to a single colony) with the following variables:
-Observation ID
-Observaton date
-Subject
-Behavior
-Modifiers
-Behavior Type
-Start 
-Stop 
-Comments


# Questions to answer

1-Do solitary queens perform different behaviors than social queens?

2-Do social queens spend proportionately more/less time on certain tasks than solitary queens?

3-Do social queens spend absolutely more/less time on certain tasks than solitary queens?


# Data Analyses

### Load Packages
```{r}
library(tidyverse)
```

### Load Data
```{r}

## create a list of all the files in the BORISfiles directory. these are all the files we want to import
borisfiles <- list.files("/Users/ericasarro/Google\ Drive\ File\ Stream/My\ Drive/WoodardLab.shared/Projects/2018.Awesome.Summer.Experiments/Raw\ Data/BORISfiles")

## loop over the files and import all of them. rename column names to remove "." and make shorter
lapply(borisfiles, read.csv, header = TRUE, col.names = c("obsID", "obsDate", "filename", "totalLen", "fps", "beeID", "behav", "mod", "behavType", "start", "stop", "dur", "comStart", "comStop"))

## do this to all the things 
do.call(rbind(filename))




##
## load a subset of data to test
all_data <- read.csv("/Users/ericasarro/Google\ Drive\ File\ Stream/My\ Drive/MasterProjects/18Su_AwesomeSummer/MockData/AggEvents.csv", header = TRUE, col.names = c("obsID", "obsDate", "filename", "totalLen", "fps", "beeID", "behav", "mod", "behavType", "start", "stop", "dur", "comStart", "comStop" ))
head(all_data)

```





### Clean Data
```{r}

## convert columns to character to enable manipulation with strsplit() and subset(startsWith())
all_data$beeID <- as.character(all_data$beeID)
all_data$obsID <- as.character(all_data$obsID)

## split out obsID into camID, nestID, dateID and hourID
all_data$camID <- sapply(strsplit(all_data$obsID, "\t"), 
                          function(x){
                            x[1]
                            } )

all_data$nestID <- sapply(strsplit(all_data$obsID, "\t"), 
                          function(x){
                            x[2]
                            } )

all_data$dateID <- sapply(strsplit(all_data$obsID, "\t"), 
                          function(x){ 
                            x[3]
                            } )

all_data$hourID <- sapply(strsplit(all_data$obsID, "\t"), 
                          function(x){ 
                            x[4]
                            } )
```

```{r}
## group data by beeID, then by behavior, then count the number of instances of each behavior

## this is grouping properly by behavior, but not separating out beeIDs
all_data %>%
  group_by(beeID) %>%
  group_by(behav) %>%
  summarise(count = sum(all_data$behav == behav))

## this for loop is only returning the last beeID, not the earlier ones
all_data$beeID <- as.factor(all_data$beeID)
for(i in 1:length(levels(all_data$beeID))) {
  beeID_data <- subset(all_data, beeID == levels(all_data$beeID)[i])
  return(beeID_data)
}

## but it properly outputs the behavior counts here for the last beeID
beeID_data %>%
  group_by(behav) %>%
  summarise(count = sum(beeID_data$behav == behav))


```

```{r}
## group data by subject
subj_data <- group_by(all_data, beeID)

## group data by behavior
behav_data <- group_by(subj_data, behav)

## calculate absolute amount of time spent on each behav
#  for each subject: sum time carrying out each task, divide by total time watched for that nestID
#  when calculating total time:
#   **need to account for videos with 0 data. make sure boris output includes those
#   **need to account for multiple events in a single video -- can only sum unique obsIDs

## for each beeID, count the number of unique behaviors
sapply(behav_data$beeID, sapply(behav_data$behav, count(behav_data$behav)))

for (i in levels(behav_data$beeID)) {
  behav_beeID <- subset(behav_data, beeID = i)
  num_feeds <- sum(behav_beeID$behav == "feedsbrood")
  return(num_feeds)
}

num_feeds


       for (all unique subjects) {
  for (all unique behaviors) {
    sum number of behav
    sum time of dur
  }
}
    

## caclulate proportion of time spent on each behav

## subset to only queens
queen_data <- subset(behav_data, startsWith(all_data$beeID, "Queen"))

```





### Visualize Data
pie charts with propotional time, bar graphs with absolute
```{r}
## saves plot
ggsave(xxxx.pdf)

```





### Analyze Data
```{r}

```





