---
title: "Queen Time Budgeting"
output: html_notebook
---

## Proposed Topic
Bumble bee queen task repertoire and time budget in response to the social environment.

## Acquired Data
Data includes several spreadsheets (each spreadsheet corresponding to a single colony) with the following variables:
-Observation ID
-Observaton date
-Subject
-Behavior
-Modifiers
-Behavior Type
-Start 
-Stop 
-Comments


### Questions to answer
Does the social environment impact bumble bee queen task repertoire?
-Do solitary queens perform different behaviors than social queens?

Does the social environment impact bumble bee queen time allocation? 
-Do social queens spend proportionally less time on certain tasks than solitary queens?
-Do social queens spend absolutely less time on certain tasks than solitary queens?





# Load Data
```{r}
## create a list of all the files in the BORISfiles directory. these are all the files we want to import
borisfiles <- list.files("/Users/ericasarro/Google\ Drive\ File\ Stream/My\ Drive/WoodardLab.shared/Projects/2018.Awesome.Summer.Experiments/Raw\ Data/BORISfiles")

## loop over the files and import all of them 
lapply(borisfiles, read.csv(header = TRUE))

## do this to all the things 
do.call(rbind(filename))



## load a subset of data to test
events <- read.csv("/Users/ericasarro/Google\ Drive\ File\ Stream/My\ Drive/MasterProjects/18Su_AwesomeSummer/MockData/AggEvents.csv", header = TRUE, col.names = c("obsID", "obsDate", "filename", "totalLen", "fps", "subject", "behav", "mod", "behavType", "start", "stop", "dur", "comStart", "comStop" ))
head(events)

```





# Clean Data
```{r}

## give a vector, will split it based on argument (ie. Space) split=“[ ]” 
Strsplit(split = "[]")
events$nestID <- sapply(strsplit(events$obsID, ""), function(x){ 
x[2]
} )

events$obsID
asdf
```





# Visualize Data
pie charts with propotional time, bar graphs with absolute
```{r}
## saves plot
ggsave(xxxx.pdf)

```





# Analyze Data
```{r}

```





